The host galaxy of the short GRB111117A at z = 2.211. \thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-05-24 GRB111117A.tex
Circulated paper

\abstract{
Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to localize. We here present evidence of the, to date, highest redshift high-confidence short GRB, located at $z=2.211$. This value supersedes a previously estimated, lower redshift value. Spectroscopic observations along with a wealth of imaging data allow us to place tight constraints on the nature of the host galaxy at an unprecedented distance. 
The rest frame X-ray derived hydrogen column density is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs). This could indicate that the part of the sGRB population hosted in late-type galaxies, live in similar environments to that of lGRB hosts. 
The host is in the brighter end of the expected host brightness distribution at $z = 2.211$, but only 30 per cent of the distribution should be missed due to host faintness at this redshift. This indicates that we are not missing redshifts for a dominant fraction of the sGRB hosts at this redshift. The redshift of GRB111117A is evidence against a lognormal delay-time model for sGRBs through the predicted redshift distribution of sGRBs, which is very sensitive to high-z sGRBs.
From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.1 $R$_\odot$ is required for the case of a binary neutron star (NS) system. This puts constraints on the progenitor system evolution up to the time of explosion.

@@

We determine a redshift of $z = 2.211$ from from the simultaneous detection of emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with \hb~detected at low significance ($\sim 3 \sigma$). \ha is only visible in the first epoch, due to the K-band blocking filter used for the remainder observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and shows significant velocity structure along the slit. A drop in the continuum the the blue of the \lya line further supports the inferred redshift.

@@

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time of the non-detection. We find $\beta_{OX} < 0.79$, consistent with what was reported in \citet{Sakamoto2013}.

Because the projected distance does not change much between redshift 1.3 and 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are unaffected.

>>>>

\abstract{
Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to localize. We here present evidence of the, to date, highest redshift high-confidence short GRB, located at $z=2.211$. This value supersedes a previously estimated, lower redshift value. Spectroscopic observations along with a wealth of imaging data allow us to place tight constraints on the nature of the host galaxy at an unprecedented distance. 
The rest frame X-ray derived hydrogen column density is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs). This could indicate that the part of the sGRB population hosted in late-type galaxies, live in similar environments to that of lGRB hosts. 
The host is in the brighter end of the expected host brightness distribution at $z = 2.211$, but only 30 per cent of the distribution should be missed due to host faintness at this redshift. This indicates that we are not missing redshifts for a dominant fraction of the sGRB hosts at this redshift. The redshift of GRB111117A is evidence against a lognormal delay-time model for sGRBs through the predicted redshift distribution of sGRBs, which is very sensitive to high-z sGRBs.
From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.2 $R$_\odot$ is required for the case of a binary neutron star (NS) system. This puts constraints on the progenitor system evolution up to the time of explosion.

@@

We determine a redshift of $z = 2.211$ from from the simultaneous detection of emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with \hb~detected at low significance ($\sim 3 \sigma$). \ha is only visible in the first epoch, due to the K-band blocking filter used for the remainder observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and shows significant velocity structure along the slit. A drop in the continuum to the blue of the \lya~line further supports the inferred redshift.

@@

%A measure of the optical-to-X-ray flux ratio is parametrized in terms of the "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time of the non-detection. We find $\beta_{OX} < 0.79$, consistent with what was reported in \citet{Sakamoto2013}.

%Because the projected distance does not change much between redshift 1.3 and 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are unaffected.
